---
title: "pipelearner"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

pipelearner makes it easy to create machine learning pipelines in R.

## Installation and background

pipelearner is currently available from github as a development package only. It can be installed by running:

```{r, eval = F}
# install.packages("devtools")
devtools::install_github("drsimonj/pipelearner")
```

pipelearner is built on top of [tidyverse](https://github.com/tidyverse/tidyverse) packages like [modelr](https://github.com/hadley/modelr). To harness the full power of pipelearner, it will help to possess some technical knowledge of tidyverse tools such as:

- `%>%` the pipe operator from [magrittr]() package.
- tibbles from [tibble]() package.
- `map()` and other iteration functions from [purrr]() package.
- `resample` objects from [modelr](https://github.com/hadley/modelr) package.

An excellent resource to get started with these is Hadley Wickham's book, [R for Data Science](http://r4ds.had.co.nz/).

## The basics

Like how ggplot2 elements are layered with `+`, pipelearner involves building a pipelearner object, which is a list, with functions that can be piped into eachother with `%>%`:

- `pipelearner()` to setup the pipelearner object
- `learn_cvpairs()` to customize the cross-validation pairs
- `learn_curves()` to customize the learning curves using incremental proportions of training data
- `learn_models()` to add new learning models.

Once the pipelearner object is setup, call `learn()` to fit everything and `summary()` to extract the results as a tibble.

## Simple case

The following sets up a single regression model `lm` for predicting `Sepal.Length` with all other variables (`Sepal.Length ~ .`) in the `iris` data set.

```{r}
library(pipelearner)

pl <- pipelearner(iris, lm, Sepal.Length ~ .)
```

Printing a pipelearner object exposes the list elements:

```{r}
pl
```

**Defaults** to note:

- `data` is split into a single cross-validation pair of resample objects (under `cv_pairs`) referencing 80% of the data for training and 20% for testing.
- Learning is done on the entire proportion of the training data (`train_ps == 1`).

When these elements as setup as desired, use `learn()` to fit all models to every combination of training proportions (`train_ps`) and set of training data in the cross-validation pairs (`cv_pairs`), and return a tibble of the results.

```{r}
pl %>% learn()
```

Quick notes:

- `fit` contains the fitted models.
- `train` contains a resample object referencing the data that each model was fitted to.
- `test` contains a resample object referencing test data that models were *not* fitted to (for later use).

## Cross-validation pairs

Cross-validation pairs can be customized with `learn_cvpairs`. The following implements k-fold cross-validation, creating five folds:

```{r}
pl %>%
  learn_cvpairs(k = 5) %>% 
  learn()
```

Notice the five rows where the model has been fitted to training data for each fold, represented by `cv_pairs.id`. The precise training data sets are also stored under `train`.

## Learning curves

Learning curves can be customized wth `learn_curves`. The following will fit the model to three proportions of the training data - .5, .75, and 1:

```{r}
pl %>% 
  learn_curves(.5, .75, 1) %>% 
  learn()
```

Notice the three rows where the model has been fitted to the three proportions of the training data, represented by `train_p`. Again, `train` contains references to the exact training data used in each case.

## More models

Add more models with `learn_models()`. For example, the following adds a decision tree to be fitted:

```{r}
pl %>% 
  learn_models(rpart::rpart, Sepal.Length ~ .) %>% 
  learn()
```

Notice two rows where the regression and decision tree models have been fit to the training data, represented by `models.id`. The different model calls also appear under `model`.

Things to know about `learn_models()`:

- It can be called multiple times within the pipeline (unlike the other `learn_*()` functions).
- It is called implicitly by `pipelearner()` when arguments beyond a data frame are supplied. For example, `pipelearner(d, l, f, ...)` is equivalent to `pipelearner(d) %>% learn_models(l, f, ...)`.
- Its arguments can all be vectors, which will be expanded to all combinations. This making it easy to do things like compare many models with the same formulas, compare many different formulas, or do grid-search.

For example, the following fits two models with three formulas:

```{r}
pipelearner(iris) %>%
  learn_models(c(lm, rpart::rpart),
               c(Sepal.Length ~ Sepal.Width,
                 Sepal.Length ~ Sepal.Width + Petal.Length,
                 Sepal.Length ~ Sepal.Width + Petal.Length + Species)) %>% 
  learn()
```

The following fits a regression model and grid-searches hyperparameters of a decision tree:

```{r}
pipelearner(iris) %>%
  learn_models(lm, Sepal.Length ~ .) %>% 
  learn_models(rpart::rpart, Sepal.Length ~ .,
               minsplit = c(2, 20), cp = c(0.01, 0.1)) %>% 
  learn()
```

## Bringing it all together

All of the functions can be combined into a single pipeline. For example, the following will:

- Create 50 cross-validation pairs (holding out random 20% of data by default in each).
- To each be fitted in proportions of .5 to 1 in increments of .1.
- With a regression modelling all interactions, and a decision tree modelling all features.

```{r}
iris %>% 
  pipelearner() %>% 
  learn_cvpairs(n = 50) %>% 
  learn_curves(seq(.5, 1, by = .1)) %>% 
  learn_models(lm, Sepal.Width ~ .*.) %>% 
  learn_models(rpart::rpart, Sepal.Width ~ .) %>% 
  learn()
```

## Extracting information from fits

As you can see, pipelearner makes it easy to fit many models. The next step is to extract performance metrics from the tibble of results. This is where prior familiarity working with tidyverse tools becomes useful, if not essential.

At present, pipelearner doesn't provide functions to extract any further information. This is because the information to be extracted can vary considerably between the models fitted to the data.

The following will demonstrate an example of visualising learning curves by extracting performance information from regression models.

`r_square()` is setup to extract an R-squared value. It is based on `modelr::rsquare`, but adjusted to handle new data sets (I've submitted an issue to incorporate into `modelr`).

```{r}
# R-Squared scoring (because modelr rsquare doen't work right now)
response_var <- function(model) {
  formula(model)[[2L]]
}
response <- function(model, data) {
  eval(response_var(model), as.data.frame(data))
}
r_square <- function(model, data) {
  actual    <- response(model, data)
  residuals <- predict(model, data) - actual
  1 - (var(residuals, na.rm = TRUE) / var(actual, na.rm = TRUE))
}
```

Using the `weather` data from the `nycflights13` package, fit a single regression model to 50 cross-validation pairs, holding out 15% of the data for testing in each case, in iterative training proportions:

```{r}
library(nycflights13)

results <- weather %>% 
  pipelearner() %>% 
  learn_cvpairs(n = 50, test = .15) %>% 
  learn_curves(seq(.1, 1, by = .1)) %>% 
  learn_models(lm, visib ~ humid + precip + wind_dir) %>% 
  learn()
results
```

New columns are added with `dplyr::mutate` containing the rsquared values for each set of training and test data by using `purrr` functions.

```{r, message = F}
library(purrr)
library(dplyr)

results <- results %>% 
  mutate(
    rsquare_train = map2_dbl(fit, train, r_square),
    rsquare_test  = map2_dbl(fit, test,  r_square)
  )

results %>% select(cv_pairs.id, train_p, contains("rsquare"))
```

We can visualize these learning curves as follows:

```{r}
library(tidyr)
library(ggplot2)

results %>% 
  select(train_p, contains("rsquare")) %>% 
  gather(source, rsquare, contains("rsquare")) %>% 
  ggplot(aes(train_p, rsquare, color = source)) +
   geom_jitter(width = .03, alpha = .3) +
   stat_summary(geom = "line", fun.y = mean) +
   stat_summary(geom = "point", fun.y = mean, size = 4) +
   labs(x = "Proportion of training data used",
        y = "R Squared")
```





